---
title: "ZV-CV test"
date: "11 Dec 2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

An example:

We plan to evaluate $E_{z\sim q}[log g(z;\theta)]dz$, where $q_t(z,\theta) = TN(\mu,\sigma/\sqrt{t},smin,smax)$ (a truncated normal distribution), $g(z;\theta)=N(\mu,\sigma)$.


We can obtain expectation analytically by:

\[ E_{z\sim q} [ log g(z;\theta)] = -0.5 log 2\pi - log\sigma - E_{z\sim q} \dfrac{(z-\mu)^2}{2\sigma^2} \]

By using the fact that $\dfrac{z-\mu}{\sigma/\sqrt{t}}$ is a standarded truncated normal distribution with boundary points $a=(smin-\mu)/(\sigma/\sqrt{t})$, $b=(smax-\mu)/(\sigma/\sqrt{t})$,  the packakage \text{truncnorm} etruncnorm(.) and vtruncnorm(.) can be used to get the value for 2nd moments.

In the implementation:

Integrand is  $log g(z;\theta)$

samples are $z$, where $z\sim TN(\mu,\sigma/\sqrt{t},smin,smax)$

derivates are $-\frac{z-\mu}{\sigma^2/t}$

Set $smin = -2$, $smax = 2$, $t = 1$ (coding for t is omitted), $\mu=0$,$\sigma=1$ and N=2000 (number of samples).

```{r}
library(ZVCV)
library(truncnorm)

true_value =function(smin,smax,mu,sigma,t){
  a = (smin - mu)/sigma*sqrt(t)
  b = (smax - mu)/sigma*sqrt(t)
  mean_value = etruncnorm(a,b,mean=0,sd=1)
  var_value = vtruncnorm(a,b,mean=0,sd=1)
  true_value = -0.5*log(2*pi)-log(sigma)- 0.5*(mean_value^2+var_value)/t
  return(true_value)
}

N <- 2000

mymean = 0
mycov = 1

require(mvtnorm)
set.seed(1)
smin_set = -2
smax_set = 2
samples <- rtruncnorm(N,a=smin_set,b=smax_set,mymean,sqrt(mycov))

integrand <- dnorm(samples,mymean,sqrt(mycov),log=TRUE)


#' # derivatives of Gaussian wrt x
derivatives <- -c(solve(mycov))*(samples-mymean)
#' 

#' # With the ZVCV package
# order 0 is the naive MC estimate
order0<-zvcv(integrand,samples,derivatives,options = list(polyorder = 0))$expectation

# MC_est <- mean(dnorm(samples,mean = mymean,sd = sqrt(mycov), log=TRUE))
# polynomial with order 1 
order1<-zvcv(integrand,samples,derivatives,options = list(polyorder = 1,regul_reg=FALSE))$expectation

# polynomial with order 2 
order2<-zvcv(integrand,samples,derivatives,options = list(polyorder = 2,regul_reg=FALSE))$expectation

# true value
tv <- true_value(smin=smin_set,smax=smax_set,mu=mymean,sigma=sqrt(mycov/1/1),t=1)


paste('smin/smax:',smin_set,'/',smax_set)

paste('true value:',tv)

paste('order 0:', order0)

paste('order 1:', order1)

paste('order 2:', order2)

```